Data Appendix to \"The impact of weather events on schooling outcomes\

### 2020-03-19 data_appendix.Rmd
*Reorganized code to match intended format. *Modified the file names to use relative paths so they will work on other machines *Switched the rename syntax to R style which can rename a whole long list of variables at once, doesn't use quotes and follows a new_name = old_name pattern. *Added notes which can be deleted later describing steps that need to be added.

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F,
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")
# input Stata file
library(foreign)
data_2005 <- read.dta("C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/Raw data/seminar_paper_data/2005_hhsurvey.dta")

@@

## Dataset description
**Citation:** .  
Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales

@@

*Once you've described the variables, enter an R chunk by selecting Code -> Insert Chunk, or Ctrl+Alt+I, give it a name to describe the dataset you are importing. After importing, export a dataframe summary using the command.*

`export_summary_table(dfSummary(dataset_name))`

*While it will make your resulting file long, you should not modify the chunk options to suppress printing of code and output. I would likely not include this in the documentation for an actual paper I was submitting, but including them here will let me read your code and the output message from R and may help identify data import concerns early in the process. Since these files will exist only electronically, their length is less of a concern. If you like to print out files to proofread and want me to help you shorten the printed versions, let me know. We can temporarily modify the chunk options for printing and restore them before you submit the assignment.*

@@

>>>>

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F,
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")
# add haven package to make stata 16 version readable 
library(foreign)
library(haven)

@@

## Household Survey data
**Citation:** .  
Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales

@@

```{r import household data}
data_2005 <- read_dta(file.path("Raw data","seminar_paper_data","2005_hhsurvey.dta")) %>% 
  rename(hh_id = folio,
        assist = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy = s3_01,
        register = s3_04,
        grade = s3_05b,
        repeated = s3_07)
# it looks like you need to add dept = xxx to the list above since dept is not showing up as a variable for me        
hh_2005 <- data_2005 %>% 
  select(hh_id, assist, reason, members, sex, age, relation, literacy, register, grade, repeated, dept)
```

```{r exporting data table}
export_summary_table(dfSummary(hh_2005))
```

## Weather data
**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
**Date Downloaded:** Identify when you downloaded the dataset.  
**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
**Unit of observation:** What distinguishes different rows in your dataset?  
**Dates covered:** What time frame does the data cover?

Describe in a step-by-step fashion how an interested user could obtain the data.

In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.

- **variable_name:** Variable description. 
- **variable_name2:** Description of second variable.

```{r import weather data}
weather <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
  rename(dept = NOM_DEP, flood = A_2005)

##Note: I can't tell completely but it appears that the file you're reading in has all the weather data. Once you load the data onto GitHub, I can help you figure out how to use pivot_longer() to reshape the weather data.

flood_summary %>% weather %>% 
  group_by(dept) %>% 
  summarise(flood = sum(flood))

`export_summary_table(dfSummary(dataset_name))`
```

@@

```{r merging datasets}
#I would recommend combining all the years of household data and then merging with the weather data

total_2005 <- hh_2005 %>% 
  left_join(flood_summary)
```

```
